---
output:
  pdf_document: default
  html_document: default
---

## Loading necessary libraries

```{r warning=FALSE}
library(ggplot2)
library(rsample)     # data splitting 
library(dplyr)       # data wrangling
library(rpart)       # performing regression trees
library(rpart.plot)  # plotting regression trees
```


## Loading the dataset

```{r}
movies_data <- read.csv("Final_Project_FlixGem.csv")
```
## Task 1

## Data Pre-Processing

Since the company doesn't want results for Series, we have to drop rows containing Series.

```{r}
movies_data_v2<-movies_data[!(movies_data$Series.or.Movie=="Series"),]
```
Moreover, we can drop rows which has missing values (NA) as well.
So our polished and cleaned data looks like this:

```{r}
movies_data_v3 <- na.omit(movies_data_v2)
```

## Part (a)

## Hidden Gem Scores and Languages

Converting the languages in string to character vectors and assigning the first 
language in the string as the language because after studying the data, the languages 
are mostly listed by abundance and so the top language is automatically selected 
as the categorical value for that movie. Here, there might be some information loss, 
but this seems like a good way to encode the data since there are just too many categories.

```{r warning=FALSE}
for (i in 1:length(movies_data_v3$Languages)) {
  movies_data_v3$Languages[i]<-unlist(strsplit(movies_data_v3$Languages[i],split = ", "))
  
}

summary_table <-movies_data_v3 %>% 
   group_by(Languages) %>% 
   summarise(n = n())

summary_table <- summary_table %>% arrange(desc(n))
knitr::kable(head(summary_table))
```
We are deciding to chose the first 5 languages and encode the rest as "Other". The 
top five languages are English, French, Spanish, Japanese and Mandarin.

```{r warning=FALSE}
important_langs <- c("English","French","Spanish","Japanese","Mandarin")

for (i in 1:length(movies_data_v3$Languages)) {
  if(!(movies_data_v3$Languages[i] %in% important_langs)){
    movies_data_v3$Languages[i]<-"Other"
  }
}

(unique(movies_data_v3$Languages))
```

These are the language categories we will be working with. Since the 
Hidden Gem Score is a continuous variable, we can use histograms to see the 
distribution of the Hidden Gem Scores for each Language.

```{r}
ggplot(movies_data_v3, aes(x=Hidden.Gem.Score,fill=Languages)) +
     geom_histogram(bins=25,col="black") +
     facet_wrap(~as.factor(Languages))+
  labs(x="Hidden Gem Score",title = "Histograms for Hidden Gem Scores based on Languages")

```

Plotting a boxplot to see the median, interquartile range and outlier values for 
the Hidden Gem Scores for each Language category.

```{r}
ggplot(movies_data_v3,aes(x=Languages,y=Hidden.Gem.Score,fill=Languages)) +
    stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
     ylab("Hidden Gem Score")+
  labs(title="Boxplots for Hidden Gem Scores for each Language")

```


Explanation

From the histrograms we can see that the majority of the movies are in English, with 
the shape of the distribution of the English movies being negatively
skewed. From the histogram, the shape of the distribution for the rest of the languages are not really clear
because of their relatively low count. From the boxplots, we see that the median Hidden Gem score for 
movies having English as one of the release-languages are the lowest while those that have Mandarin
are the highest. The median Hidden Gem score for the rest of the languages are relatively the same (around 3.7).
From the boxplot, we also see that English, Spanish and the 'other' categoires have a negative skew while
movies in Mandarin, and French have a postive skew. Finally, the ditribution for the movies in Japanese
is symmetric. We also note that moveis supporting English has the most number of 
outliers and those in Mandarin have the lowest.


## Hidden Gem Scores and Runtime

Since the Hidden Gem Score is a continuous variable, we can use histograms to see the distribution of the Hidden Gem Scores for each different Runtime Category to see if there is any relationship between them:

Plotting a histogram for each runtime category to see distribution of Hidden Gem Scores.
```{r}
ggplot(movies_data_v3, aes(x=Hidden.Gem.Score,fill=Runtime)) +
     geom_histogram(bins=25,col="black") +
     facet_wrap(~as.factor(Runtime))+
  labs(x="Hidden Gem Score",title = "Histograms for Hidden Gem Scores for each Runtime")

```

Plotting a boxplot to see the median, interquartile range and outlier values for the Hidden Gem Scores for each different Runtime Category.
```{r}
ggplot(movies_data_v3,aes(x=Runtime,y=Hidden.Gem.Score,fill=Runtime)) +
    stat_boxplot(geom="errorbar",width=0.25) + geom_boxplot() +
     ylab("Hidden Gem Score")+
  labs(title="Boxplots for Hidden Gem Scores for each Runtime")

```


Explanation

From the boxlpots we see that for all three categories of run time the median and 
the upper quartile values are relatively the same, although the lower quartile 
values vary quite a bit. From both the histogram and the boxplots, we see that 
the less than 30 mins and 1-2 hour run time category all have the same negative
skew. From the hsitrogram, we see that the shape of the distribution for 
the greater than 2 hour category and the 1-2 hour category are relatively the same.
Therefore, we can say the greater than 2 hours category of run time also has the same negative skew.


## Part (b)

## IMDb Scores and Hidden Gem Scores

Analyzing if IMDb Scores are correlated with Hidden Gem Scores. Plotting a graph with IMDb Scores on the x-axis and Hidden Gem Scores on the y-axis.

```{r}
ggplot(movies_data_v3,aes(x=IMDb.Score,y=Hidden.Gem.Score)) + geom_point() +
xlab("IMBb Score") + ylab("Hidden Gem Score")
```

From the graph, we can see that there is some positive correlation between IMDb 
Scores and Hidden Gem Scores since the graph has a positive gradient except some 
randomly scattered Hidden Gem Score values(around 5 to 8) when the IMDb Scores 
vary between 6 and 8.


## Rotten Tomatoes Scores and Hidden Gem Scores

Analyzing if Rotten Tomatoes Scores are correlated with Hidden Gem Scores. Plotting 
a graph with Rotten Tomatoes Scores on the x-axis and Hidden Gem Scores on the y-axis.

```{r}
ggplot(movies_data_v3,aes(x=Rotten.Tomatoes.Score,y=Hidden.Gem.Score)) + geom_point() +
xlab("Rotten Tomatoes Score") + ylab("Hidden Gem Score")
```

From the graph, we can deduce that there is a strong positive correlation between 
Rotten Tomatoes Score and Hidden Gem Score since the graph has a positive gradient 
since it is upward sloping. However, there are a few movies within high rotten 
tomatoes scores which have much higher hidden gem scores in proportion.

## Metacritic Scores and Hidden Gem Scores

Analyzing if Metacritic Scores are correlated with Hidden Gem Scores. Plotting a 
graph with Metacritic Scores on the x-axis and Hidden Gem Scores on the y-axis.

```{r}
ggplot(movies_data_v3,aes(x=Metacritic.Score,y=Hidden.Gem.Score)) + geom_point() +
xlab("Metacritic Score") + ylab("Hidden Gem Score")
```

Again, it looks like there is a positive correlation due to the upward sloping 
nature of the scatterplot. This suggests there is strong positive correlation.



## Part(c)

From the boxplpots we can see that the median values of the Hidden Gem scores don't
change much as the length of the runtime increase. Also, from the histograms for the 
1-2 hour category and the >2 hour category, the shape of the distribution remains relatively 
same. Therefore, we can't really claim that people are becoming more acceptable 
of longer movies.

\newpage

## Task 2

Creating the regression tree with outcome as Hidden Gem Score and predictors as 
Language, IMDb Score, Runtime, Rotten Tomatoes Score and Metacritic Score

```{r}
regression_tree <- rpart(
  Hidden.Gem.Score ~ Languages + Runtime + IMDb.Score+ Rotten.Tomatoes.Score + 
    Metacritic.Score,
  data = movies_data_v3, 
  method = "anova"
)

regression_tree
```


## Plotting the Regression Tree 

```{r}
rpart.plot(regression_tree)
```

Summary

Since the root of our tree is the Rotten Tomatoes score we can say that the 
Rotten Tomatoes Score is the most important in determining the Hidden Gem Score. 
We could also say that languages are the second most important factor followed by 
IMDb scores in determining the said score. The larger the number of leaves,
the more accurate our predictions would be. We believe that our model has sufficient
number of leaves, given the range of Hidden Gem scores in the data set, 
to predict the score with relative accuracy. We also note that if we had mores
leaves in our tree our predictions would be more accurate.

\newpage

## Task 3

Creating empty data frame to store the Director Names and the HG-H Index.

```{r}
director_data<-data.frame( matrix(ncol = 2, nrow = length(unique(movies_data_v3$Director))))
colnames(director_data) <- c("Director","HGHIndex")
director_data$Director <- unique(movies_data_v3$Director)
knitr::kable(head(director_data))
```

Gathering the Hidden Gem Scores for each director

```{r}

find_hgs <- function(director){
  scores <- c()
  for(j in 1:length(movies_data_v3$Director)){
    if(movies_data_v3$Director[j]==director){
      #take hidden gem scores
      scores <- append(scores,movies_data_v3$Hidden.Gem.Score[j])
    }
  }
  return(scores)
}
#function for finding the h_index
h_index <- function(gem_scores) {
    sorted_gemscores <- sort(gem_scores, decreasing = T)
    bools<- (sum(sorted_gemscores >= seq_along(sorted_gemscores)))
    return(bools)
}

for (i in 1:length(director_data$Director)) {
  director_data$HGHIndex[i]<-h_index(find_hgs(director_data$Director[i]))
}

knitr::kable(head(director_data))

```

## Top 10 Directors

```{r}
director_data<-director_data %>% arrange(desc(HGHIndex))
knitr::kable(director_data[1:10,])
```
Note that there are more directors with the same Hidden Gem Score H-Index of 4 but they are not displayed here.













